---
title: "USB results"
output: html_document
date: "2023-08-23"
---
Clear the R enviorment
```{r}
rm(list = ls())
```


## Packages
Import all the relevant packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(dplyr)
library(tidyr)
library(kableExtra)
library(stringr)
library(rstan)
library(patchwork)
library(gtools)
library(posterior)
library(cmdstanr)
seed <- 20200331
set.seed(seed)
```

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/home/teodor/Desktop/projects/papers/journal1/code/')
```

```{r cars}
source("utils.R")
```

```{r}
dataset <- readr::read_csv("results/full_data_noise.csv") %>% dplyr::filter(task=="text")

dataset %>% sample_n(10) %>% kable(
    "latex", 
    table.envir = 'table',
    caption='Sample of the generated dataset', 
    booktabs=T,
    label='dataset_sample',
    format.args = list(scientific = FALSE), 
    digits = 3,
    linesep = "") %>% 
    kable_styling(latex_options = c("hold_position"),
                  full_width = F) 
```

```{r}

bm <- get_index_names_as_array(dataset$dataset)
algorithms <- get_index_names_as_array(dataset$algorithm)
```

## 

```{r}
y <- ifelse(dataset$accuracy>90,1,0)

dataset<-dataset %>% 
  dplyr::mutate(y=y) %>%
  dplyr::group_by(algorithm, dataset, SD) %>%
  dplyr::summarise(
    solvedAt90=sum(y),
    N=n()) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(algorithmID=create_index(algorithm),
                datasetID=create_index(dataset)) %>%
  dplyr::select(algorithm, algorithmID, dataset, datasetID, SD, N, y=solvedAt90)
```

Lets preview a sample of the data set
```{r}
kable(dplyr::sample_n(dataset,size=10),"html", booktabs=T, format.args = list(scientific = FALSE), digits = 3) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>% 
  kableExtra::scroll_box(width = "100%")
```


```{r include=F}
saveRDS(dplyr::sample_n(dataset,size=4),'./probsuccessmodeldata.RDS')
```

```{r}
standata <- list(
  N_total = nrow(dataset),
  y = dataset$y,
  N_draw = dataset$N,
  x_noise = dataset$SD,
  N_algorithm = length(algorithms),
  algorithm_id = dataset$algorithmID,
  N_bm  = length(bm),
  bm_id = dataset$datasetID
)
saveRDS(standata,file="./probsuccess-data.RDS")
```

```{r}
standata<-readRDS("./probsuccess-data.RDS")

cmdstanr::set_cmdstan_path("/home/teodor/.cmdstan/cmdstan-2.35.0")
file <- file.path(cmdstanr::cmdstan_path(), "probsuccess-cv.stan")
mod <- cmdstanr::cmdstan_model(file)

probsuccess.fit <- mod$sample(
  data = standata,
  seed = 123,
  chains = 4,
  iter_warmup = 200,
  iter_sampling = 2000,
  parallel_chains = 4,
  refresh = 500
)

probsuccess.fit$save_object(file = "./probsuccess-fit-text.RDS")

```

```{r}
probsuccess.fit <-readRDS("./probsuccess-fit-text.RDS")
np <- bayesplot::nuts_params(probsuccess.fit)
a_alg <- c("a_alg[1]",
           "a_alg[2]",
           "a_alg[3]",
           "a_alg[4]",
           "a_alg[5]",
           "a_alg[6]",
           "a_alg[7]",
           "a_alg[8]",
           "a_alg[9]",
           "a_alg[10]",
           "a_alg[11]",
           "a_alg[12]",
           "a_alg[13]")
b_noise <- c("b_noise[1]",
             "b_noise[2]",
             "b_noise[3]",
             "b_noise[4]",
             "b_noise[5]",
             "b_noise[6]",
             "b_noise[7]",
             "b_noise[8]",
             "b_noise[9]",
             "b_noise[10]",
             "b_noise[11]",
             "b_noise[12]",
             "b_noise[13]")
a_bm_norm <- c("a_bm_norm[1]",
               "a_bm_norm[2]",
               "a_bm_norm[3]",
               "a_bm_norm[4]"
               )

csv_files <- probsuccess.fit$output_files()
#probsuccess.fit <- as_cmdstan_fit(csv_files)
draws_a <- posterior::as_draws(probsuccess.fit$draws(variables = a_alg, format = "draws_df"))
draws_a <- draws_a[c(1000:1999, 3000:3999, 5000:5999, 7000:7999),]

draws_b <- posterior::as_draws(probsuccess.fit$draws(variables = b_noise, format = "draws_df"))
draws_b <- draws_b[c(1000:1999, 3000:3999, 5000:5999, 7000:7999),]

draws_a_bm <- posterior::as_draws(probsuccess.fit$draws(variables = a_bm_norm, format = "draws_df"))
draws_a_bm <- draws_a_bm[c(1000:1999, 3000:3999, 5000:5999, 7000:7999),]
```
Another diagnosis is to look at the Rhat. If Rhat is greater than 1.05 it indicates a divergence in the chains (they did not mix well). The table below shows a summary of the sampling.
```{r}
summary_probsuccess.fit <- probsuccess.fit$summary(variables = a_alg) %>%
  dplyr::select(-mean, -median,-sd,-mad,-q5,-q95 )%>% 
  kable(
      "latex", 
      table.envir = 'table',
      caption='Diagnostics for the posterior effects (text data)', 
      booktabs=T,
      label='tab:diagnostics-text-probsuccess',
      format.args = list(scientific = FALSE), 
      digits = 3,
      linesep = "") %>% 
  kable_styling(latex_options = c("hold_position"),
                full_width = F) %>% 
  readr::write_lines('diagnostics-probsuccess-text.tex')
```

Here we are expimenting to compute the HPDintervals
```{r}
hpdi <- coda::HPDinterval(coda::as.mcmc(t(as.data.frame(probsuccess.fit$summary(variables = c(a_alg,b_noise,a_bm_norm))[6:7]))))
df <- tibble::rownames_to_column(as.data.frame(hpdi), "Parameter")
Mean <- c(probsuccess.fit$summary(variables = c(a_alg,b_noise,a_bm_norm))$mean)
df.hpdi <- mutate(df,
                 Mean=Mean,
                 Parameter=probsuccess.fit$summary(variables = c(a_alg,b_noise,a_bm_norm))$variable) %>%
  rename(HPDI.lower=lower,
         HPDI.higher=upper,
         Mean = Mean)
hpdi <- df.hpdi
hpdi_oddsratio <- df.hpdi
hpdi_oddsratio$Mean <- exp(df.hpdi$Mean)
hpdi_oddsratio$HPDI.lower <- exp(df.hpdi$HPDI.lower)
hpdi_oddsratio$HPDI.higher <- exp(df.hpdi$HPDI.higher)
```

Produce tables.
```{r}
rename_pars <- c(
  paste(rep('a_',length(algorithms)),algorithms, sep = ""),
  paste(rep('b_',length(algorithms)),algorithms, sep = ""))

renamepars <- rename_pars
pars <- c(a_alg, b_noise)
stanfit_summary <- as_tibble(probsuccess.fit$summary(variables = c(a_alg,b_noise)), rownames="Parameter")
t1 <- stanfit_summary %>%
  dplyr::filter(Parameter %in% pars) %>%
  select(Parameter)
t2 <- hpdi %>%
  dplyr::filter(Parameter %in% pars)
t <-left_join(x=t2, y=t1, by=c("Parameter")) %>%
  dplyr::select(Parameter,Mean, everything()) %>%
  dplyr::mutate(Parameter=renamepars)
tt<- t %>% 
  mutate('OR HPD low' = exp(HPDI.lower),
         'OR Mean' = exp(Mean),
         'OR HPD high' = exp(HPDI.higher))
colnames(tt) <- c("Parameter", "Mean", "HPD low", "HPD high",'OR HPD low','OR Mean','OR HPD high')

 saveRDS(tt,'./probsuccess-par-table-text.RDS')
```

Create table for the ODDs ratios
```{r}
tt <-readRDS('./probsuccess-par-table-text.RDS')
tt[,c(1,6)] %>%
kable(
        "latex", 
        table.envir = 'table',
        caption='Odds ratios for fixed effects and noise parameters (text data)', 
        booktabs=T,
        label='summarystatisticstable-OR-text',
        format.args = list(scientific = FALSE), 
        digits = 3,
        linesep = "") %>% 
    kable_styling(latex_options = c("hold_position"),
                  full_width = F) %>% 
    readr::write_lines('summarystatisticstable-OR-text.tex')
```

Statistical Analysis
```{r}
p_trace <- bayesplot::mcmc_trace(draws_a, pars = a_alg)
png("traceplots-probsuccess-text.png")
p_trace
```

Scatterplots for MCMC
```{r}
# p_mcmc_scatter <- bayesplot::mcmc_scatter(probsuccess.fit$draws(variables=c("a_alg[1]","a_alg[2]"),format = "draws_df"))
# p_mcmc_scatter
```

```{r}
# p_mcmc_intervals<-bayesplot::mcmc_intervals(draws_a) +
#   scale_y_discrete(labels=algorithms)+
#   labs(x="Estimate",
#        y="Algorithm",
#        title="HPD Intervals for the parameters")
# png("strenght-text.png")
# p_mcmc_intervals
```
 Sample from the posterior
```{r}
y <- standata$y
y_rep <- posterior::as_draws(probsuccess.fit$draws(variables = "y_rep", format = "draws_df"))
y_rep <- y_rep[,1:length(y)]

y_sim <- posterior::as_draws(probsuccess.fit$draws(variables = "y_sim", format = "draws_df"))
y_sim <- y_rep[,1:length(y)]
```
Kernel density plots
```{r}
# p_dense_overlay <- bayesplot::ppc_dens_overlay(y,as.matrix(y_rep))
# png("dense_overlay-text.png")
```


Histograms of skew(y_rep)
```{r}
# p_bars_stat <- bayesplot::ppc_stat(y, as.matrix(y_rep), binwidth=0.005)
# png("stat-text.png")
# p_bars_stat
```

```{r}
# p_areas <- bayesplot::mcmc_areas(draws_a, prob=0.8)
# png("posterior-parameters-areas.png")
# p_areas
```

```{r}
p_bars_stat_grouped <- bayesplot::ppc_stat_grouped(y=y, yrep=as.matrix(y_rep), group = dataset$algorithm, binwidth = 0.005)
png("posterior-parameters-grouped-text-probsuccess.png")
p_bars_stat_grouped
```
Posterior Predictive Check
```{r}
p_bars_grouped <- bayesplot::ppc_stat_grouped(y=y, yrep=as.matrix(y_rep), group=dataset$algorithm, binwidth = 0.005)
#pdf('posterior-predictive-check-bars_grouped-aggregated.pdf')
png("posterior-parameters-grouped-probsuccess-text.png")
p_bars_grouped
```
Prior Predictive Check
```{r}
p_bars_grouped_prior <- bayesplot::ppc_stat_grouped(y=y, yrep=as.matrix(y_sim), group=dataset$algorithm, binwidth = 0.005)
#pdf('posterior-predictive-check-bars_grouped-aggregated.pdf')
png("prior-parameters-grouped-probsuccess-text.png")
p_bars_grouped_prior
```

```{r}
a_posterior <- hpdi$Mean[1:13]
b_noise_posterior <- hpdi$Mean[14:26]
a_bm_posterior <- hpdi$Mean[27:38]

N_total = nrow(dataset)
p <- rep(1,N_total)
algorithm_ID <- rep(1,N_total)
dataset_ID <- rep(1,N_total)

algorithm <-  rep(1,N_total)
dataset_name <-  rep(1,N_total)
noise <- rep(1,N_total)
task <- rep(1,N_total)

for (j in 1:N_total) {
  p[j] <- inv.logit(a_posterior[dataset$algorithmID[j]]+a_bm_posterior[dataset$datasetID[j]]+b_noise_posterior[dataset$algorithmID[j]]*dataset$SD[j])
  algorithm[j] <- dataset$algorithm[j]
  dataset_name[j] <- dataset$dataset[j]
  
  algorithm_ID[j] <- dataset$algorithmID[j]
  dataset_ID[j] <- dataset$datasetID[j]
  noise[j] <- dataset$SD[j]
}

probability <- tibble(p=as.numeric(p),algorithm,algorithm_ID,dataset_name,dataset_ID,noise) %>% 
  select(p,algorithm,noise) %>%
  group_by(algorithm, noise)

summary_prob <- probability %>%
  summarize(avg_prob = 100*mean(p, na.rm = TRUE))

summary_prob_0 <- summary_prob %>%
  dplyr::filter(noise==0) %>%
  dplyr::select(-noise) %>%
  dplyr::rename("SD=0 (text)"=avg_prob)

summary_prob_1 <- summary_prob %>%
  dplyr::filter(noise==3) %>%
  dplyr::select(-noise) %>%
  dplyr::rename("SD=1 (text)"=avg_prob)

summary_probability <- dplyr::left_join(summary_prob_0,summary_prob_1,by=join_by(algorithm))
saveRDS(summary_probability,'./probability-of-success-text.RDS')

```




