---
title: "redo-journal1"
output: html_document
date: "2024-06-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Clear the R enviorment
```{r}
rm(list = ls())
```


## Packages
Import all the relevant packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(dplyr)
library(tidyr)
library(kableExtra)
library(stringr)
library(rstan)
library(patchwork)
library(gtools)
library(posterior)
library(cmdstanr)
library(dplyr)
library(ggplot2)
seed <- 20200331
set.seed(seed)
```

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = '/home/teodor/Desktop/projects/papers/journal1/code/')
```

```{r cars}
source("utils.R")
```

## Import dataset
```{r}
dataset <- readr::read_csv("results/full_data.csv") %>%
  dplyr::filter(task=="numeric")

dataset %>% sample_n(10) %>% kable(
    "latex", 
    table.envir = 'table',
    caption='Sample of the generated dataset', 
    booktabs=T,
    label='dataset_sample',
    format.args = list(scientific = FALSE), 
    digits = 3,
    linesep = "") %>% 
    kable_styling(latex_options = c("hold_position"),
                  full_width = F)
```
## Summary statistics for the error rate
```{r}
dataset%>% 
  dplyr::group_by(algorithm) %>% 
  summarise("Median" = median(accuracy),
            "5%" = quantile(accuracy, 0.05),
            "95%" = quantile(accuracy, 0.95),
            "Range" = quantile(accuracy, 0.95) - quantile(accuracy, 0.05)) %>% 
  dplyr::ungroup() %>% 
  kable(
        "latex", 
        table.envir = 'table',
        caption='Summary statistics for the error rate (numeric data)', 
        booktabs=T,
        label='numeric',
        format.args = list(scientific = FALSE), 
        digits = 3,
        linesep = "") %>% 
    kable_styling(latex_options = c("hold_position"),
                  full_width = F) %>% 
    readr::write_lines('summarystatisticstable-accuracy-numeric.tex')
```

## Summary statistics but as boxplots
```{r}
p<-ggplot(data=dataset, aes(x=algorithm, y=accuracy, fill=algorithm))+
  geom_boxplot()+
  theme(axis.text.x = element_blank())+ #remove the x labels
  labs(title = 'accuracy (all datasets)')
pdf("boxplots-accuracy-numeric.pdf")
#png("boxplots-accuracy-numeric.png")
p

```
## Widen dataset, in addition we remove the task column
```{r}
data_wide <- dataset %>% pivot_wider(names_from = dataset, values_from = accuracy)
```

```{r}
dataset %>% 
  dplyr::group_by(algorithm) %>% 
  summarise("Median" = median(accuracy),
            "5%" = quantile(accuracy, 0.05),
            "95%" = quantile(accuracy, 0.95),
            "Range" = quantile(accuracy, 0.95) - quantile(accuracy, 0.05)) %>% 
  dplyr::ungroup() %>% 
  kable(
        "latex", 
        table.envir = 'table',
        caption='Summary statistics for the accuracy (numeric data)', 
        booktabs=T,
        label='summarystatisticstable-numeric',
        format.args = list(scientific = FALSE), 
        digits = 3,
        linesep = "") %>% 
    kable_styling(latex_options = c("hold_position"),
                  full_width = F) %>% 
    readr::write_lines('summarystatisticstable-numeric.tex')
```
## Prepare data for analysis
```{r}
dagg <- dataset %>% 
  dplyr::filter(seed<=3) %>% # Minimial amount of successful trials
  #dplyr::select(-`...1`) %>%
  dplyr::group_by(seed) %>%
  dplyr::mutate(Rank=rank(accuracy, ties.method = "random")) %>%
  dplyr::ungroup() %>%
  dplyr::select(-accuracy) #%>%
  #tidyr::pivot_wider(names_from = algorithm,
  #                   values_from = Rank)

dagg <- dagg %>% pivot_wider(names_from = algorithm, values_from = Rank) %>% drop_na()
```

```{r}
algorithms <- get_index_names_as_array(dataset$algorithm)
n_algorithms <- length(algorithms)
comb <- gtools::combinations(n=n_algorithms, r=2, v=seq(1:n_algorithms), repeats.allowed = F)
```

```{r}
pb <- progress::progress_bar$new(format = "[:bar] :current/:total (:percent)", total = nrow(dagg))

df_out <-  dplyr::tribble(~algo0_name, ~algo0, ~algo1_name, ~algo1, ~y, ~seed, ~dataset)

for(i in 1:nrow(dagg))
{
  current_row <- dagg[i,]
  for(j in 1:nrow(comb)){
    comb_row <- comb[j,]
    
    algo0_name <- algorithms[comb_row[1]]
    algo0 <- comb_row[1]
    algo0_rank <- current_row[[1,algo0_name]]
    
    algo1_name <- algorithms[comb_row[2]]
    algo1 <- comb_row[2]
    algo1_rank <- current_row[[1,algo1_name]]
    
    diff_rank <- algo1_rank - algo0_rank
    y <- ifelse(diff_rank<=0, 1, 0) 
    
    df_out <- add_row(df_out,
                      algo0_name=algo0_name,
                      algo0=algo0,
                      algo1_name=algo1_name,
                      algo1=algo1,
                      y=y,
                      seed=current_row$seed,
                      dataset=current_row$dataset)
    
  }
  pb$tick()
}
saveRDS(df_out, file="./ranking-numeric.RDS")
```

```{r pressure, echo=FALSE}
df_out <- readRDS("./ranking-numeric.RDS")
df_out$numLabId <- create_index(df_out$dataset)
benchmarks <- get_index_names_as_array(df_out$dataset)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```{r}
df_table<-df_out %>% select(-seed)
```
Create the standata
```{r}
standata <- list(
  N_total=nrow(df_out),
  y = as.integer(df_out$y),
  N_algorithm = as.integer(n_algorithms),
  algo0=as.integer(df_out$algo0),
  algo1=as.integer(df_out$algo1),
  bm_id=df_out$numLabId,
  N_bm=length(benchmarks)
)
saveRDS(standata, file = "./rankingmodel-data-numeric.RDS")
```
CmdStanR
```{r}
standata<-readRDS("./rankingmodel-data-numeric.RDS")

library(cmdstanr)
cmdstanr::set_cmdstan_path("/home/teodor/.cmdstan/cmdstan-2.35.0")
file <- file.path(cmdstanr::cmdstan_path(), "rankingmodel_withcluster.stan")
mod <- cmdstanr::cmdstan_model(file)

ranking.fit <- mod$sample(
  data = standata,
  seed = 123,
  chains = 4,
  iter_warmup = 200,
  iter_sampling = 2000,
  parallel_chains = 4,
  refresh = 500
)
saveRDS(ranking.fit, file = "./ranking-fit-numeric.RDS")
ranking.fit$save_object(file = "./ranking-fit-numeric.RDS")

```
Read in the fitted data
```{r}
ranking.fit <-readRDS("./ranking-fit-numeric.RDS")
np <- bayesplot::nuts_params(ranking.fit)
a_alg <- c("a_alg[1]",
           "a_alg[2]",
           "a_alg[3]",
           "a_alg[4]",
           "a_alg[5]",
           "a_alg[6]",
           "a_alg[7]",
           "a_alg[8]",
           "a_alg[9]",
           "a_alg[10]",
           "a_alg[11]",
           "a_alg[12]",
           "a_alg[13]"
)
draws_a <- posterior::as_draws(ranking.fit$draws(variables = a_alg, format = "draws_df"))
```

Summary statistics
```{r}
options(pillar.sigfig = 3)
summary_ranking.fit <- ranking.fit$summary(variables = a_alg) %>%
  dplyr::rename(Variable = variable, Rhat = rhat) %>%
  dplyr::select(-mean, -median,-sd,-mad,-q5,-q95 )%>% 
  kable(
      "latex", 
      table.envir = 'table',
      caption='Diagnostics for the posterior ranks (numeric data)', 
      booktabs=T,
      label='tab:diagnostics-numeric',
      format.args = list(scientific = FALSE), 
      digits = 3,
      linesep = "") %>% 
  kable_styling(latex_options = c("hold_position"),
                full_width = F) %>% 
  readr::write_lines('diagnostics-numeric.tex')
```

Statistical Analysis
```{r}

p_trace <- bayesplot::mcmc_trace(draws_a, pars=a_alg)
#pdf("traceplots-10.pdf")
png("traceplots-numeric.png")
p_trace
```

Scatterplots
```{r}
#p_mcmc_scatter <- bayesplot::mcmc_scatter(ranking.fit$draws(variables = c("a_alg[1]","a_alg[2]"),format = "draws_df"))
```

Scatterplots mcmc_pairs instead of mcmc:scatter.
```{r}
# draws_a_1 <- posterior::as_draws(ranking.fit$draws(variables = c("a_alg[1]","a_alg[2]","a_alg[3]"), format = "draws_df"))
# draws_a_2 <- posterior::as_draws(ranking.fit$draws(variables = c("a_alg[4]","a_alg[5]","a_alg[6]"), format = "draws_df"))
# draws_a_3 <- posterior::as_draws(ranking.fit$draws(variables = c("a_alg[7]","a_alg[8]","a_alg[9]"), format = "draws_df"))
# draws_a_4 <- posterior::as_draws(ranking.fit$draws(variables = c("a_alg[10]","a_alg[11]","a_alg[12]"), format = "draws_df"))
# draws_a_5 <- posterior::as_draws(ranking.fit$draws(variables = c("a_alg[12]","a_alg[13]"), format = "draws_df"))
# 
# 
# 
# 
# p_mcmc_pairs_1 <- bayesplot::mcmc_pairs(draws_a_1, np=np)
# p_mcmc_pairs_2 <- bayesplot::mcmc_pairs(draws_a_2, np=np)
# p_mcmc_pairs_3 <- bayesplot::mcmc_pairs(draws_a_3, np=np)
# p_mcmc_pairs_4 <- bayesplot::mcmc_pairs(draws_a_4, np=np)
# p_mcmc_pairs_5 <- bayesplot::mcmc_pairs(draws_a_5, np=np)
# pdf("scatter-numeric.pdf")
# 
# 
# p_mcmc_pairs_1
# p_mcmc_pairs_2
# p_mcmc_pairs_3
# p_mcmc_pairs_4
# p_mcmc_pairs_5
```

```{r}
y <- df_out$y
y_rep <- posterior::as_draws(ranking.fit$draws(variables = "y_rep", format = "draws_df"))
y_rep <- y_rep[,1:length(y)]

y_sim <- posterior::as_draws(ranking.fit$draws(variables = "y_sim", format = "draws_df"))
y_sim <- y_sim[,1:length(y)]
#library(shinystan)
#launch_shinystan(ranking.fit)
```

Parplots
```{r}
# p_parcoord <- bayesplot::mcmc_parcoord(draws_a)
# #pdf("parcoord-numeric.pdf")
# png("parcoord-numeric.png")
# p_parcoord
```
Kernel density estimates
```{r}
# p_dense_overlay <-bayesplot::ppc_dens_overlay(y,as.matrix(y_rep))
# #pdf("dense_overlay-numeric.pdf")
# png("dense_overlay-numeric.png")
# p_dense_overlay
```
Hisograms of skew(y_rep)
```{r}
# p_ppc_stat <- bayesplot::ppc_stat(y, as.matrix(y_rep), binwidth = 0.005)
# #pdf('stat-numeric.pdf')
# png("stat-numeric.png")
# p_ppc_stat
```
## Results

```{r}
p_mcmc_intervals<-bayesplot::mcmc_intervals(draws_a) +
  scale_y_discrete(labels=algorithms)+
  labs(x='Estimate',
       y='Algorithm',
       title='Strength parameters')
#pdf('strength-numeric.pdf')
png("strength-numeric.png")
p_mcmc_intervals
```
POSTERIOR PREDICTIVE CHECKS

Posterior distribution of a_alg
```{r}
draws_a <- draws_a %>% dplyr::rename(centeredkernel=`a_alg[1]`,
                          laplace=`a_alg[2]`,
                          "mean_shifted_laplace"=`a_alg[3]`,
                          poisson=`a_alg[4]`,
                          poisson2=a_alg[5],
                          poissonbalanced=a_alg[6],
                          poissonmbo=`a_alg[7]`,
                          poissonmbo_old=`a_alg[8]`,
                          poissonmbobalanced=a_alg[9],
                          poissonvolume=`a_alg[10]`,
                          randomwalk=`a_alg[11]`,
                          sparselabelpropagation=`a_alg[12]`,
                          wnll=`a_alg[13]`
                          )
p_areas<-bayesplot::mcmc_areas(draws_a, prob = 0.8)
#pdf('posterior-strength-parameters-numeric.pdf')
png("posterior-strength-parameters-numeric.png")
p_areas
```

```{r}
# p_bars <- bayesplot::ppc_bars(y=y,yrep=as.matrix(y_rep))
# #pdf('posterior-predictive-check-boxplot-numeric.pdf')
# png("posterior-predictive-check-boxplot-numeric.png")
# p_bars
```

```{r}
p_bars_grouped <- bayesplot::ppc_stat_grouped(y=y, yrep=as.matrix(y_rep), group=df_out$algo0_name, binwidth = 0.005)
png("posterior-parameters-grouped-numeric.png")

p_bars_grouped

```

```{r}
p_bars_grouped_prior <- bayesplot::ppc_stat_grouped(y=y, yrep=as.matrix(y_sim), group=df_out$algo0_name, binwidth = 0.005)
png("prior-parameters-grouped-numeric.png")
p_bars_grouped_prior

```

````{r}
# p_rootogram <- bayesplot::ppc_rootogram(y=y,yrep=as.matrix(y_rep))
# #pdf('posterior-predictive-check-rootogram-numeric.pdf')
# png("posterior-predictive-check-rootogran.numeric.png")
# p_rootogram

```
 


Here we are extracting all samples and ranking them to have a distribution of the ranks
```{r}
posterior_df <- as.data.frame(posterior::as_draws_df(ranking.fit$draws(variables = a_alg)))[,1:n_algorithms]
colnames(posterior_df) <- algorithms
#sampling from the posterior
s <- dplyr::sample_n(posterior_df, size = 1000, replace=T)
s <- dplyr::mutate(s, rown = row_number())
wide_s <- tidyr::pivot_longer(s, cols=all_of(algorithms), names_to = "Algorithms", values_to = "a_alg")

rank_df <- wide_s %>% 
  dplyr::group_by(rown) %>% 
  dplyr::mutate(Rank = rank(-a_alg, ties.method = 'random')) %>% 
  dplyr::ungroup() %>% 
  dplyr::select(-a_alg) %>% 
  dplyr::group_by(Algorithms) %>% 
  dplyr::summarise(MedianRank = median(Rank),
                   VarianceRank = var(Rank)) %>% 
  dplyr::arrange(MedianRank)
```

```{r}
# rank_df_table <- rank_df
# colnames(rank_df_table) <- c("Models","Median Rank", "Variance of the Rank")
# kable(rank_df_table, "html") %>% 
#   kable_styling(bootstrap_options = c('striped',"hover", "condensed" ))
```
saving the table to latex
```{r}
rank_df_table %>% 
  kable(
      "latex", 
      table.envir = 'table',
      caption='Ranking of the algorithms (numeric data)', 
      booktabs=T,
      label='rankingnumeric',
      format.args = list(scientific = FALSE), 
      digits = 3,
      linesep = "") %>% 
  kable_styling(latex_options = c("hold_position"),
                full_width = F) %>% 
  readr::write_lines('ranking-numeric.tex')
```
